Trends from 30-Year Observations of Downward Solar Irradiance in Thessaloniki, \colorbox{green}{Greece}

### 2023-12-27 MDPI_submission.Rmd
Commit 2023-12-27 11:10

Due to the significant spatial and temporal variability of the trends and the contributing factors, there is a constant need to monitor and investigate SDR at different sites in order to estimate the degree of variability, and its relation to the local conditions.
In this study, we examine the trends of SDR, using ground-based measurements at Thessaloniki, Greece, for the period from `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`.
<!-- goals -->
We re-evaluated and extended the dataset used by @Bais2013, we applied a different algorithm for the identification of clear-/cloud-sky instances [@Reno2016; @Reno2012a], and we derived the SDR trends for the period of study, under different sky conditions (all-sky, clear-sky, and cloudy-sky).
Finally, we investigated the dependence of the trends on solar zenith angle and season.

>>>>

Due to the significant spatial and temporal variability of the trends and the contributing factors, there is a constant need to monitor and investigate SDR at different sites in order to estimate the degree of variability, and its relation to the local conditions.
In this study, we examine the trends of SDR, using ground-based measurements at Thessaloniki, Greece, for the period from `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`.
<!-- goals -->
We re-evaluated and extended the dataset used by @Bais2013, we applied a different algorithm for the identification of clear-/cloud-sky instances [@Reno2016; @Reno2012], and we derived the SDR trends for the period of study, under different sky conditions (all-sky, clear-sky, and cloudy-sky).
Finally, we investigated the dependence of the trends on solar zenith angle and season.

### 2023-12-27 MDPI_submission.tex
Commit 2023-12-27 11:10

There is a consensus among researchers that the major factor affecting
the variability of SDR attenuation is the interactions of solar
radiation with atmospheric aerosols and clouds. Those interactions,
among other factors, have been analyzed with models
\citep{Li2016, Samset2018}, showing the existence of feedback mechanisms
between the two. Similar findings have been shown from the analysis of
observations at other locations
\citep{Schwarz2020, Ohvril2009, Zerefos2009, Xia2007} {[}and references
therein{]}.\\
In the Mediterranean region aerosols have been recognized as an
important factor affecting the penetration of solar radiation at the
surface \citep{Fountoulakis2016, Siomos2018, Gkikas2013, Lozano2021}.
These studies investigated the long-term trend in aerosol optical depth,
which has been found to decrease in the last three decades, the
transport and composition of aerosols, and their radiative effects.

Due to the significant spatial and temporal variability of the trends
and the contributing factors, there is a constant need to monitor and
investigate SDR at different sites in order to estimate the degree of
variability, and its relation to the local conditions. In this study, we
examine the trends of SDR, using ground-based measurements at
Thessaloniki, Greece, for the period from 1993 to 2023. We re-evaluated
and extended the dataset used by \citet{Bais2013}, we applied a
different algorithm for the identification of clear-/cloud-sky instances
\citep{Reno2016, Reno2012a}, and we derived the SDR trends for the
period of study, under different sky conditions (all-sky, clear-sky, and
cloudy-sky). Finally, we investigated the dependence of the trends on
solar zenith angle and season.

>>>>

There is a consensus among researchers that the major factor affecting
the variability of SDR attenuation is the interactions of solar
radiation with atmospheric aerosols and clouds. Those interactions,
among other factors, have been analyzed with models
\citep{Li2016, Samset2018}, showing the existence of feedback mechanisms
between the two. Similar findings have been shown from the analysis of
observations at other locations
\citep{Schwarz2020, Ohvril2009, Zerefos2009, Xia2007} {[}and references
therein{]}. In the Mediterranean region aerosols have been recognized as
an important factor affecting the penetration of solar radiation at the
surface \citep{Fountoulakis2016, Siomos2018, Gkikas2013, Lozano2021}.
These studies investigated the long-term trend in aerosol optical depth,
which has been found to decrease in the last three decades, the
transport and composition of aerosols, and their radiative effects.

Due to the significant spatial and temporal variability of the trends
and the contributing factors, there is a constant need to monitor and
investigate SDR at different sites in order to estimate the degree of
variability, and its relation to the local conditions. In this study, we
examine the trends of SDR, using ground-based measurements at
Thessaloniki, Greece, for the period from 1993 to 2023. We re-evaluated
and extended the dataset used by \citet{Bais2013}, we applied a
different algorithm for the identification of clear-/cloud-sky instances
\citep{Reno2016, Reno2012}, and we derived the SDR trends for the period
of study, under different sky conditions (all-sky, clear-sky, and
cloudy-sky). Finally, we investigated the dependence of the trends on
solar zenith angle and season.

### 2023-12-27 manuscript.v5/applsci-2743235.tex
Commit 2023-12-27 11:10

Due to the significant spatial and temporal variability of the trends
and the contributing factors, there is a constant need to monitor and
investigate SDR at different sites in order to estimate the degree of
variability and its relation to the local conditions. In this study, we
examine the trends of SDR using ground-based measurements at
Thessaloniki, Greece, for the period from 1993 to 2023. We re-evaluated and
extended the dataset used by \citet{Bais2013}, we applied a different
algorithm for the identification of clear-/cloud-sky instances
\citep{Reno2016, Reno2012a}, and we derived the SDR trends for the
period of study under different sky conditions (all-sky, clear-sky, and
cloudy-sky). Finally, we investigated the dependence of the trends on
solar zenith angle and season.

@@

    \end{adjustwidth}
 \caption{Time series of \colorbox{red}{monthly} %MDPI: Please add the explanation for subfifgures in the figure.  %Natsis: Done that, and also rephrased the caption
 mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots). The black lines represent the linear trends, (\textbf{a}-\textbf{c}) winter, (\textbf{d}-\textbf{f}) spring, (\textbf{g}-\textbf{i}) summer, (\textbf{j}-\textbf{i}) autumn, and also (\textbf{a}, \textbf{d}, \textbf{g}, \textbf{i}) all-skies, (\textbf{b}, \textbf{e}, \textbf{h}, \textbf{k}) clear-skies and (\textbf{c}, \textbf{f}, \textbf{i}, \textbf{j}) cloudy-skies.}\label{fig:seasonalALL}
\end{figure}

@@

% Only for the journal Methods and Protocols:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title. A supporting video article is available at doi: link.}
\authorcontributions{\hl{  } %MDPI: For research articles with several authors, a short paragraph specifying their individual contributions must be provided. The following statements should be used ``Conceptualization, X.X. and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y. and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing---original draft preparation, X.X.; writing---review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y. All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation. Authorship must be limited to those who have contributed substantially to the work~reported.
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\institutionalreview{\hl{  } %MDPI: In this section, you should add the Institutional Review Board Statement and approval number, if relevant to your study. You might choose to exclude this statement if the study did not require ethical approval. Please note that the Editorial Office might ask you for further information. Please add “The study was conducted in accordance with the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving humans. OR “The animal study protocol was approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving animals. OR “Ethical review and approval were waived for this study due to REASON (please provide a detailed justification).” OR “Not applicable” for studies not involving humans or animals.
}

@@

\begin{figure}[H]
    \begin{adjustwidth}{-\extralength}{0cm}
        {\centering
            \includegraphics[width=1.0\linewidth]{./images/SzaTrendsSeasTogether-2}
        }
   \end{adjustwidth}
        \caption{\colorbox{red}{Long} %MDPI: Please add the explanation for subfifgures in the figure.
 -term trends of SDR as a function of SZA separately form morning
               and afternoon periods, by season (rows of plots) for the three sky
               conditions (columns of plots).
               Solid shapes represent statistical significant trends ($p<0.005$).
               Cases where $p<0.005$ or with less than $85$~observations may be
               missing from view.}\label{fig:SZAtrendSeason}

>>>>

Due to the significant spatial and temporal variability of the trends
and the contributing factors, there is a constant need to monitor and
investigate SDR at different sites in order to estimate the degree of
variability and its relation to the local conditions. In this study, we
examine the trends of SDR using ground-based measurements at
Thessaloniki, Greece, for the period from 1993 to 2023. We re-evaluated and
extended the dataset used by \citet{Bais2013}, we applied a different
algorithm for the identification of clear-/cloud-sky instances
\citep{Reno2016, Reno2012}, %Natsis: citation edited
and we derived the SDR trends for the
period of study under different sky conditions (all-sky, clear-sky, and
cloudy-sky). Finally, we investigated the dependence of the trends on
solar zenith angle and season.

@@

    \end{adjustwidth}
 \caption{Time series of \colorbox{red}{monthly} %MDPI: Please add the explanation for subfifgures in the figure.  %Natsis: Done that, and also rephrased the caption
 mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots). The black lines represent the linear trends, (\textbf{a}-\textbf{c}) winter, (\textbf{d}-\textbf{f}) spring, (\textbf{g}-\textbf{i}) summer, (\textbf{j}-\textbf{l}) autumn, and also (\textbf{a}, \textbf{d}, \textbf{g}, \textbf{j}) all-skies, (\textbf{b}, \textbf{e}, \textbf{h}, \textbf{k}) clear-skies and (\textbf{c}, \textbf{f}, \textbf{i}, \textbf{l}) cloudy-skies.}\label{fig:seasonalALL}
\end{figure}

@@

% Only for the journal Methods and Protocols:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title. A supporting video article is available at doi: link.}
\authorcontributions{\hl{  } %MDPI: For research articles with several authors, a short paragraph specifying their individual contributions must be provided. The following statements should be used ``Conceptualization, X.X. and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y. and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing---original draft preparation, X.X.; writing---review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y. All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation. Authorship must be limited to those who have contributed substantially to the work~reported.
Conceptualization, methodology, software, data analysis and original writing---draft preparation A.N.; writing---review and editing, supervision, A.B.; data curation, C.M. All authors have read and agreed to the published version of the manuscript.} %Natsis: section added
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\institutionalreview{\hl{  } %MDPI: In this section, you should add the Institutional Review Board Statement and approval number, if relevant to your study. You might choose to exclude this statement if the study did not require ethical approval. Please note that the Editorial Office might ask you for further information. Please add “The study was conducted in accordance with the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving humans. OR “The animal study protocol was approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving animals. OR “Ethical review and approval were waived for this study due to REASON (please provide a detailed justification).” OR “Not applicable” for studies not involving humans or animals.
Not applicable}  %Natsis: section added

@@

\begin{figure}[H]
    \begin{adjustwidth}{-\extralength}{0cm}
        {\centering
            \includegraphics[width=1.0\linewidth]{./images/SzaTrendsSeasTogether-2}
        }
   \end{adjustwidth}
        \caption{\colorbox{red}{Long}%MDPI: Please add the explanation for subfifgures in the figure.
 -term trends of SDR as a function of SZA separately form morning
               and afternoon periods, by season (rows of plots) for the three sky
               conditions (columns of plots).
               Solid shapes represent statistical significant trends ($p<0.005$), (\textbf{a}-\textbf{c}) winter, (\textbf{d}-\textbf{f}) spring, (\textbf{g}-\textbf{i}) summer, (\textbf{j}-\textbf{l}) autumn, and also (\textbf{a}, \textbf{d}, \textbf{g}, \textbf{j}) all-skies, (\textbf{b}, \textbf{e}, \textbf{h}, \textbf{k}) clear-skies and (\textbf{c}, \textbf{f}, \textbf{i}, \textbf{l}) cloudy-skies.
               Cases where $p<0.005$ or with less than $85$~observations may be
               missing from view.}\label{fig:SZAtrendSeason}
